Modeling the Linux page cache for accurate simulation of data-intensive applications

### 2020-12-10 paper_io_sim.tex
Pass on the "simulation" part of the Related work section Added many bib references

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{paralist}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{ulem}
\usepackage{tabularx}

@@

@@

        % Some cosmetic rewrites below
    In this paper, we propose a I/O simulation model that characterizes
    the key features of the page cache. We have implemented this model
    as  part of the WRENCH workflow simulation framework, which itself
    builds on the popular SimGrid distributed systems simulation
    framework.  Our model and its implementation enable the simulation
    of both single-threaded and multithreaded applications, and of both
    writeback and writethrough caches for local or network-based
    filesystems. We evaluate the accuracy of our model in different
    conditions, including sequential and concurrent applications, as
    well as local and remote I/Os.  We find that our page cache model
    reduces the simulation error by up to an order of magnitude when
    compared to state-of-the-art, cacheless simulations.

@@

        % Swapped the order the two paragraphs and merged them into one
        The number of data-intensive applications has been steadily rising as a result of
        open-data and data sharing initiatives. Due to the sheer size of the data being
        processed, these applications must be executed on large scale infrastructure
    such as High Performance Computing (HPC) clusters or the cloud.  It
    is thus crucial to quantify the performance of these applications
    on these platforms. The goals include determining which type of hardware/software
    stacks are best suited to different application classes, as well as
    understanding the shortcomings of current algorithms, designs and
    technologies so as to propose new solutions. Unfortunately, application performance
        studies via experiments on real-world compute platforms
        faces several difficulties (high operational costs, labor-intensive experimental setups,
        shared platforms with dynamic loads that hinder reproductibilty of results) and shortcomings
        (experiments are limited to the available platform/softare configurations, which precludes
        the exploration of hypothetical scenarios). 
    Simulations address these concerns by providing simulation models
    and abstractions for the performance of computer hardware, such as
    CPU, network and storage. Consequently, simulations provide a
    cost-effective, fast, easy and reproducible way to evaluate
    application performance on arbitrary platform configurations. It thus comes
        as no surprise that, considering here only the field of parallel and distributed
        computing, a large number of simulation frameworks have been developed and used
        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.

@@

    In this work, we present WRENCH-cache, a simulation model of I/O
    with page cache incorporated into the WRENCH simulation framework~\cite{casanova2020fgcs},
    which is  based on the popular SimGrid distributed simulation framework~\cite{casanova2014simgrid}.
    The added value of WRENCH-cache is to make it possible to simulate 
        application execution accurately in the presence of page caching. Our specific 
        contributions are as follows:
    \begin{compactitem}
        \item A page cache simulation model that supports 
    both single-threaded and multithreaded applications, and of both
    writeback and writethrough caches for local or network-based
    filesystems;
        \item An implementation of this model in WRENCH; 
        \item An evaluation of the accuracy and scalability of our model, and of its implementation,
              for a range of applications, platforms, and page cache configurations. 
    \end{compactitem}

@@

            Many simulation frameworks exist to describe the behaviour of applications
            running on the various infrastructures available to users, such as personal computers,
            clouds and clusters.
            These simulators are built with resource models such CPU, network and
            storage, using various different approaches to describe the hardware.
            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models
            and \textit{DiskSim} the simulates storage hardware functioning \tristan{add references}.

            Simulators must choose between scalability and accuracy when describing the
            hardware~\cite{casanova2014simgrid}.
            Fine-grained models allow simulators to achieve high accuracy,
            but lead to scalability issues due to long simulation time
            when simulated platforms and applications grow.
            SimGrid is an open source simulation platform that has been developed
            not only to balance the trade-off between accuracy and scalability,
            but also to achieve versatility, allowing the use of the framework
            to simulate applications in multiple domains. WRENCH, a workflow management
            system simulator based on the SimGrid framework, provides a high-level API
            for simulating workflows, thus facilitating the integration of new models.
            \valerie{say more about wrench}

            Despite the fact that many sophisticated models for resources have been
            integrated in simulators, and the page cache shows apparent impacts on I/O
            performance, the simulation model of page cache is still missing in most
            of simulators.
            For example, SimGrid only simulates I/O operations with storage bandwidths
            and capacity.
            SIMCAN is a simulation framework that models page cache by storing data
            accessed on disk in a block cache \cite{nunez2012simcan}.
            Page cache is also modeled in the iCanCloud simulator through a component which
            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
            the scalability of the iCanCloud simulator is limited by its sophisticated model.
            Ultimately, none of these simulators provide any writeback cache simulator nor
            cache eviction policys through LRU lists.
            Although cache replacement policies are applied in \cite{xu2018saving} to simulate
            in-memory caching, their simulator is specific to energy consumption of multi-tier
            heterogenous networks.

            In this study, we use WRENCH
            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
            from its SimGrid backend, as well as its high-level API.
            We have integrated our model in WRENCH and use the SimGrid API to interact with
            simulated storage resources.
            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of
            resource consumption.
            We also leverage the existing SimGrid storage model with capacity
            and bandwidth properties, which are similar to the characteristics of memory.

@@

    At any given time, a file can be partially cached, completely cached,
    or uncached at all. A cached data block can only reside in one of two
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to the top of the active list. Cached blocks
    written to cache are marked as dirty until they are flushed to disk.

@@

            To validate our simulation model, we created a simple prototype
            simulator independent of existing simulation frameworks and libraries.
            This allowed us to evaluate the accuracy and correctness of our
            model in a simple scenario before integrating it to the more complex
            WRENCH framework.
            In this prototype we used the following basic storage model for
            both memory and disk:
            \begin{align*}
                & t_{r} = D / b_r \\
                & t_{w} = D / b_w\
            \end{align*}

@@

            We also implemented our model as part of WRENCH, enhancing its internal implementation
            and APIs with a page cache abstraction, and allowing users to activate the feature via a
            command-line argument. For the experiments in this paper we use WRENCH 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
            SimGrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}). However, our implementation is now part of WRENCH's master branch and will be available to users with the next WRENCH release.

\tristan{anything else interesting to mention about the implementation?}
\henri{I mentioned that the implementaion is now part of the WRENCH master }

@@

        Our experiments compare real executions with our Python prototype,
        with the original WRENCH, and with our WRENCH-cache extension. We
        evaluate our page cache simulation in single-threaded and
        multi-threaded applications, accessing data on local and remote
        file systems.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since SimGrid, and thus WRENCH, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|l
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Bandwidths  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

            As our focus is on I/O rather than compute, we measured
            application task durations on a cluster node and used these
            durations in our simulation. For the Python simulator,
            execution times were directly used in the simulation. For
            WRENCH and WRENCH-cache, we determined the corresponding number
            of flops on a 1~Gflops CPU and used these values in the
            simulation (see Table~\ref{table:cputime}). The final simulated
            platform and application used in our experiment are avaiable at
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

@@

            % Finally, because WRENCH simulates applications base on network-communication,
            % we use an infinite bandwidth to eliminate network latency for local I/Os.

@@

            Finally, we simulate a real application from the neuroimaging
            domain. \textcolor{red}{[Description of the real pipeline with
            nighres (including the workflow engine if it's Dask)]}.
            We simulate this pipeline with simulators using original WRENCH
            and WRENCH-cache. Because our work focuses on I/O time, we
            assume that CPU time is correctly modeled and use the CPU time
            measured in the real pipelines to setup our simulated
            application.

@@

        The page cache simulation model drastically reduces I/O simulation
        errors in the three successive application tasks (Figure~\ref{fig:single_error}). The first read is not impacted
        as it only involves uncached data. Errors are reduced from an average
        of 345\% in the original WRENCH to 46\% in the Python prototype and
        39\% in WRENCH-cache. Unsurprisingly, the original WRENCH simulator
        significantly overestimates read and write times, due to its lack
        of page cache simulation model. Results with files of 50~GB and 75~GBs
        show similar behaviors and are not reported due to space limitations.

        WRENCH simulation errors are substantially lower with 100~GB
        files than with 20~GB files, due to the fact that part of the
        100~GB file needs to be read and written to disk, the only storage
        device in WRENCH, as it does not fit in cache. Conversely,
        simulation errors of the Python prototype and WRENCH-cache are higher with
        100~GB files than with 20~GB files, due to idiosyncracies in the kernel
        flushing and eviction strategies.

        Simulated memory profiles are highly consistent with the real ones
        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory
        profiles almost exactly match the real ones, although dirty data
        seems to be flushing faster in real life than in simulation
        \tristan{any idea why?}, a behavior also observed with 100~GB files. With
        100~GB files, used memory reaches total memory during the first write,
        triggering dirty data flushing. Used memory drops to cached memory
        when application tasks release anonymous memory. Simulated cached
        memory is highly consistent with real values, except toward the end
        of Read 3 where it slightly increases in simulation but not in
        reality \tristan{why?}. As expected, dirty data remains under the
        dirty ratio. The Python prototype and WRENCH-cache exhibit nearly
        identical memory profiles, which reinforces the confidence in our
        implementation.

@@

            The page cache model notably reduce WRENCH's simulation error
            for concurrent applications and local I/Os
            (Figure~\ref{fig:multi_local}). For reads, WRENCH-cache still
            slightly overestimates reality, due to the discrepancy between
            simulated and real read bandwidths mentioned before
            (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes,
            WRENCH-cache slightly underestimates reality for the same reason. The plateau
            ending around 7 concurrent pipelines in the WRENCH-cache curve
            marks the limit beyond which the page cache is saturated with dirty data
            and needs flushing \tristan{why is there no plateay irl?}.

@@

            Page cache simulation importantly reduces the simulation error
            on remote storage as well (Figure~\ref{fig:multi_nfs}). This
            manifests only for reads, as only write-through cache was
            enabled in the NFS server. Both WRENCH and WRENCH-cache
            underestimate write times due to the discrepancy between
            simulated and real bandwidths mentioned previously. For reads,
            this discrepancy only impacts the results beyond 22
            applications since before this threshold, most reads result in cache
            hits \tristan{Dzung, pls check that.}.

@@

        \subsection{Overhead comparison}
        As is the case for WRENCH, simulation time with WRENCH-cache scales
        linearly with the number of concurrent applications
        (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
        cache model substantially increases the simulation time by
        application as can be seen by comparing regression slopes in
        Figure~\ref{fig:multi_time}. Interestingly, WRENCH-cache is faster with 
        NFS I/Os than with local I/Os, most likely due to the use of writethrough
        cache in NFS, which by-passes flushing operations.
        \begin{figure}
            \begin{subfigure}{\columnwidth}
                \centering
                \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
            \end{subfigure}
            \caption{Simulation time comparison. WRENCH-cache scales
            linearly with the number of concurrent applications, albeit
            with a higher overhead than WRENCH.}
            \label{fig:multi_time}
            \end{figure}

    \section{Conclusion}
    \label{discussion}
        Page cache undoubtedly has important impacts on performance, in
        particular for data-intensive applications. Numerical simulation is
        an effective approach to predict and compare the performance of
        applications, schedulers and infrastructures. However, page cache
        is missing in many simulation tools for distributed systems. In
        this study, we proposed a simulation model of I/Os with page cache,
        we implemented it in the SimGrid-based WRENCH library, and we
        evaluated its benefits in different conditions. Results show that
        our model very substantially improves simulation accuracy,
        decreasing absolute relative errors between simulated and real
        performance by a factor close to 10 (see results of the
        single-threaded experiment). Furthermore, this level of accuracy
        can be achieved without a fine-grained level of details
        \tristan{not sure what you mean}. Our page cache model is available
        in WRENCH from commit \tristan{abcd}, ready to be used for the
        simulation of Big Data applications.

        Simulation results could be made even more accurate by a deeper
        investigation of cache eviction mechanisms \tristan{be more
        specific}. The availability of asymmetrical bandwidths in the
        forthcoming SimGrid release will also directly improve our
        simulation accuracy. Relevant future work includes simulating
        application memory consumption, random I/O, file readahead and
        persistent disk storage.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{paralist}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{tabularx}

@@

\newcommand{\simgrid}{SimGrid\xspace}
\newcommand{\wrench}{WRENCH\xspace}

@@

        % Some cosmetic rewrites below
    In this paper, we propose a I/O simulation model that characterizes
    the key features of the page cache. We have implemented this model
    as  part of the \wrench workflow simulation framework, which itself
    builds on the popular \simgrid distributed systems simulation
    framework.  Our model and its implementation enable the simulation
    of both single-threaded and multithreaded applications, and of both
    writeback and writethrough caches for local or network-based
    filesystems. We evaluate the accuracy of our model in different
    conditions, including sequential and concurrent applications, as
    well as local and remote I/Os.  We find that our page cache model
    reduces the simulation error by up to an order of magnitude when
    compared to state-of-the-art, cacheless simulations.

@@

        % Swapped the order the two paragraphs and merged them into one
        The number of data-intensive applications has been steadily rising as a result of
        open-data and data sharing initiatives. Due to the sheer size of the data being
        processed, these applications must be executed on large scale infrastructure
    such as High Performance Computing (HPC) clusters or the cloud.  It
    is thus crucial to quantify the performance of these applications
    on these platforms. The goals include determining which type of hardware/software
    stacks are best suited to different application classes, as well as
    understanding the shortcomings of current algorithms, designs and
    technologies so as to propose new solutions. Unfortunately, application performance
        studies via experiments on real-world compute platforms
        faces several difficulties (high operational costs, labor-intensive experimental setups,
        shared platforms with dynamic loads that hinder reproducibility of results) and shortcomings
        (experiments are limited to the available platform/software configurations, which precludes
        the exploration of hypothetical scenarios). 
    Simulations address these concerns by providing simulation models
    and abstractions for the performance of computer hardware, such as
    CPU, network and storage. Consequently, simulations provide a
    cost-effective, fast, easy and reproducible way to evaluate
    application performance on arbitrary platform configurations. It thus comes
        as no surprise that, considering here only the field of parallel and distributed
        computing, a large number of simulation frameworks have been developed and used
        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.

@@

    In this work, we present \wrench-cache, a simulation model of I/O
    with page cache incorporated into the \wrench simulation framework~\cite{casanova2020fgcs},
    which is  based on the popular \simgrid distributed simulation framework~\cite{casanova2014simgrid}.
    The added value of \wrench-cache is to make it possible to simulate 
        application execution accurately in the presence of page caching. Our specific 
        contributions are as follows:
    \begin{compactitem}
        \item A page cache simulation model that supports 
    both single-threaded and multithreaded applications, and of both
    writeback and writethrough caches for local or network-based
    filesystems;
        \item An implementation of this model in \wrench; 
        \item An evaluation of the accuracy and scalability of our model, and of its implementation,
              for a range of applications, platforms, and page cache configurations. 
    \end{compactitem}

@@

	    Many simulation frameworks have been developed to enable the
	    simulation of parallel and distributed
	    applications~\cite{optorsim, gridsim, groudsim, cloudsim,
	    nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf,
	    cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid,
	    ROSS, casanova2020fgcs}. These frameworks implement simulation
	    models and abstractions to aid the development of simulators
	    for studying the functional and performance behaviors of
	    application workloads executed on various hardware/software
	    infrastructures.

	    The two main concerns for simulation are \emph{accuracy}
	    (the ability to faithfully reproduce real-world executions) and
	    \emph{scalability}  (the ability to simulate large/long real-world
	    executions quickly and with low RAM footprint), and the above
	    frameworks achieve different compromises between the two.  At
	    one extreme are discrete-event models that capture
	    ``microscopic'' behaviors of hardware/software systems (e.g.,
	    packet-level network simulation, block-level disk simulation,
	    cycle-accurate CPU simulation), which favor accuracy over
	    speed.  At the other extreme are analytical models that capture
	    ``macroscopic'' behaviors via mathematical models.  While these
	    models lead to fast simulation, they must be developed
	    carefully if high levels of accuracy are to be
	    achieved~\cite{velhoTOMACS2013}.

	    In this  work, we use the \simgrid and \wrench simulation
	    frameworks.  The years of research and development invested in
	    the popular \simgrid simulation framework~\cite{casanova2014simgrid}, have
	    culminated in a set of state-of-the-art macroscopic simulation
	    models that yield high accuracy, as demonstrated by
	    (in)validation studies and comparisons to competing
	    frameworks~\cite{smpi_validity, velhoTOMACS2013, simutool_09,
	    nstools_07, lebre2015, pouilloux:hal-01197274,
	    smpi_tpds2017,  7885814, 8048921, 7384330}.  But one
	    significant drawback of \simgrid is that its simulation
	    abstractions are low-level, meaning that implementing a
	    simulator of complex systems can be
	    labor-intensive~\cite{kecskemeti_2014}. To remedy this problem,
	    the \wrench simulation framework~\cite{casanova2020fgcs}
	    builds on top of \simgrid to provide higher-level simulation
	    abstractions, so that simulators of complex applications and
	    systems can be implemented with a few hundred lines.

	    Although the Linux page cache has a large impact on I/O
	    performance, and thus on the execution of data-intensive
	    applications, its simulation is rarely considered in the above
	    frameworks.  Most frameworks merely simulate I/O operations
	    based on storage bandwidths and capacities.  The SIMCAN
	    framework does models page caching by storing data accessed on
	    disk in a block cache~\cite{nunez2012simcan}.  Page cache is
	    also modeled in iCanCloud through a component that manages
	    memory accesses and cached data~\cite{nunez2012icancloud}.
	    However, the scalability of the iCanCloud simulator is limited
	    due to it using  microscopic simulation models.  Besides, none
	    of these simulators provide any writeback cache simulator nor
	    cache eviction policies through LRU lists.  Although cache
	    replacement policies are applied in~\cite{xu2018saving} to
	    simulate in-memory caching, their simulator is specific to
	    energy consumption of multi-tier heterogeneous networks.

            In this study, we develop a page cache simulation model which we implement
            as part of the \wrench framework. We targeted \wrench because it is a
            recent, actively developed framework that provides convenient simulation
            abstractions, because it is easy to incorporate new simulation models
            into it, and because it builds on \simgrid, whose scalable and accurate
            simulation models we can thus reuse for our purpose.

@@

    At any given time, a file can be partially cached, completely cached,
    or not cached at all. A cached data block can only reside in one of two
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to the top of the active list. Cached blocks
    written to cache are marked as dirty until they are flushed to disk.

@@

            To validate our simulation model, we created a simple prototype
            simulator independent of existing simulation frameworks and libraries.
            This allowed us to evaluate the accuracy and correctness of our
            model in a simple scenario before integrating it to the more complex
            \wrench framework.
            In this prototype we used the following basic storage model for
            both memory and disk:
            \begin{align*}
                & t_{r} = D / b_r \\
                & t_{w} = D / b_w\
            \end{align*}

@@

            We also implemented our model as part of \wrench, enhancing its internal implementation
            and APIs with a page cache abstraction, and allowing users to activate the feature via a
            command-line argument. For the experiments in this paper we use \wrench 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
            \simgrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}). However, our implementation is now part of \wrench's master branch and will be available to users with the next \wrench release.

\tristan{anything else interesting to mention about the implementation?}
\henri{I mentioned that the implementaion is now part of the \wrench master }

@@

        Our experiments compare real executions with our Python prototype,
        with the original \wrench, and with our \wrench-cache extension. We
        evaluate our page cache simulation in single-threaded and
        multi-threaded applications, accessing data on local and remote
        file systems.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|l
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Bandwidths  & Cluster (real) & Python simulator & \wrench simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

            As our focus is on I/O rather than compute, we measured
            application task durations on a cluster node and used these
            durations in our simulation. For the Python simulator,
            execution times were directly used in the simulation. For
            \wrench and \wrench-cache, we determined the corresponding number
            of flops on a 1~Gflops CPU and used these values in the
            simulation (see Table~\ref{table:cputime}). The final simulated
            platform and application used in our experiment are available at
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

@@

            % Finally, because \wrench simulates applications base on network-communication,
            % we use an infinite bandwidth to eliminate network latency for local I/Os.

@@

            Finally, we simulate a real application from the neuroimaging
            domain. \textcolor{red}{[Description of the real pipeline with
            nighres (including the workflow engine if it's Dask)]}.
            We simulate this pipeline with simulators using original \wrench
            and \wrench-cache. Because our work focuses on I/O time, we
            assume that CPU time is correctly modeled and use the CPU time
            measured in the real pipelines to setup our simulated
            application.

@@

        The page cache simulation model drastically reduces I/O simulation
        errors in the three successive application tasks (Figure~\ref{fig:single_error}). The first read is not impacted
        as it only involves uncached data. Errors are reduced from an average
        of 345\% in the original \wrench to 46\% in the Python prototype and
        39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
        significantly overestimates read and write times, due to its lack
        of page cache simulation model. Results with files of 50~GB and 75~GBs
        show similar behaviors and are not reported due to space limitations.

        \wrench simulation errors are substantially lower with 100~GB
        files than with 20~GB files, due to the fact that part of the
        100~GB file needs to be read and written to disk, the only storage
        device in \wrench, as it does not fit in cache. Conversely,
        simulation errors of the Python prototype and \wrench-cache are higher with
        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
        flushing and eviction strategies.

        Simulated memory profiles are highly consistent with the real ones
        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory
        profiles almost exactly match the real ones, although dirty data
        seems to be flushing faster in real life than in simulation
        \tristan{any idea why?}, a behavior also observed with 100~GB files. With
        100~GB files, used memory reaches total memory during the first write,
        triggering dirty data flushing. Used memory drops to cached memory
        when application tasks release anonymous memory. Simulated cached
        memory is highly consistent with real values, except toward the end
        of Read 3 where it slightly increases in simulation but not in
        reality \tristan{why?}. As expected, dirty data remains under the
        dirty ratio. The Python prototype and \wrench-cache exhibit nearly
        identical memory profiles, which reinforces the confidence in our
        implementation.

@@

            The page cache model notably reduce \wrench's simulation error
            for concurrent applications and local I/Os
            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
            slightly overestimates reality, due to the discrepancy between
            simulated and real read bandwidths mentioned before
            (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes,
            \wrench-cache slightly underestimates reality for the same reason. The plateau
            ending around 7 concurrent pipelines in the \wrench-cache curve
            marks the limit beyond which the page cache is saturated with dirty data
            and needs flushing \tristan{why is there no plateay irl?}.

@@

            Page cache simulation importantly reduces the simulation error
            on remote storage as well (Figure~\ref{fig:multi_nfs}). This
            manifests only for reads, as only write-through cache was
            enabled in the NFS server. Both \wrench and \wrench-cache
            underestimate write times due to the discrepancy between
            simulated and real bandwidths mentioned previously. For reads,
            this discrepancy only impacts the results beyond 22
            applications since before this threshold, most reads result in cache
            hits \tristan{Dzung, pls check that.}.

@@

        \subsection{Overhead comparison}
        As is the case for \wrench, simulation time with \wrench-cache scales
        linearly with the number of concurrent applications
        (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
        cache model substantially increases the simulation time by
        application as can be seen by comparing regression slopes in
        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
        NFS I/Os than with local I/Os, most likely due to the use of writethrough
        cache in NFS, which by-passes flushing operations.
        \begin{figure}
            \begin{subfigure}{\columnwidth}
                \centering
                \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
            \end{subfigure}
            \caption{Simulation time comparison. \wrench-cache scales
            linearly with the number of concurrent applications, albeit
            with a higher overhead than \wrench.}
            \label{fig:multi_time}
            \end{figure}

    \section{Conclusion}
    \label{discussion}
        Page cache undoubtedly has important impacts on performance, in
        particular for data-intensive applications. Numerical simulation is
        an effective approach to predict and compare the performance of
        applications, schedulers and infrastructures. However, page cache
        is missing in many simulation tools for distributed systems. In
        this study, we proposed a simulation model of I/Os with page cache,
        we implemented it in the \simgrid-based \wrench library, and we
        evaluated its benefits in different conditions. Results show that
        our model very substantially improves simulation accuracy,
        decreasing absolute relative errors between simulated and real
        performance by a factor close to 10 (see results of the
        single-threaded experiment). Furthermore, this level of accuracy
        can be achieved without a fine-grained level of details
        \tristan{not sure what you mean}. Our page cache model is available
        in \wrench from commit \tristan{abcd}, ready to be used for the
        simulation of Big Data applications.

        Simulation results could be made even more accurate by a deeper
        investigation of cache eviction mechanisms \tristan{be more
        specific}. The availability of asymmetrical bandwidths in the
        forthcoming \simgrid release will also directly improve our
        simulation accuracy. Relevant future work includes simulating
        application memory consumption, random I/O, file readahead and
        persistent disk storage.
